BlueTracer: \\ a Robust API Tracer for Evasive Malware

### 2018-07-14 evaluation.tex
Thesis grammar fixes

In order to assess BlueTracer's effectiveness with respect to real malware instances, we tested it with five highly-evasive samples collected by Joe Security \cite{JoeBox} having unusual evasion tricks that their specialized sandboxing products can take care of. For each sample, the evaluation of BlueTracer's tracing and stealth abilities was conducted by collecting BlueTracer's logs and checking if they were congruous with the sample's actions uncovered by Joe Sandbox, a commercial state-of-the-art stealthy analysis system \cite{JoeSandbox}. Moreover, we used Process Monitor, adequately cloaked by BluePill, as ground truth for what was carried out in the system.
We ran each sample for one hour using BlueTracer's \textbf{Main Image} mode.

@@

From the obtained results, we could deduce that BlueTracer was successful at tracing the sample's actions while remaining hidden and allowing them to perform their original malicious operations. An overview of the evasive actions executed from the sample's main executable which were tracked by BlueTracer and disabled by BluePill can be observed from \textit{Table 4.4} and \textit{Table 4.5}. In particular, the eight categories in which the actions are grouped correspond the ones adopted by BluePill. Specifically, the evasive patterns are classified as follows:

@@

In addition to counting the sample's events from their logs, we also recorded how many events were traced for each of eight different categories, namely, file activities, registry activities, process activities, thread activities, time activities, memory activities, system information-retrieval activities and user interface activities (\textit{Table 4.7}). Such groups are the same as the ones in Joe Sandbox's reports, which were, in fact, employed to determine the appropriate category for each event. From the data in \textit{Table 4.7} we can see that, for \textbf{1} and \textbf{3}, the most common operations are file activities (7.66\% and 59.68\%), coherently with the fact that one of their main actions is to drop files on the infected system. Regarding \textbf{2}, the largest portion of events involved thread activities (10.94\%) and memory activities (9.28\%), as we are dealing with a persistent dropper. Olympic, being a data-wiping malware, also mainly performs memory activities (9.44\%). Lastly, even for CCleaner, the majority of events are memory-related (19.84\%).

\section{Conclusions}
In this chapter we have outlined the results of BlueTracer's experimental evaluation. In order to assess BlueTracer's run-time overhead, we have tested it on a number of applications performing various tasks, employing a four different modes of operation. We have also introduced the overhead per event metric as a way of quantifying the run-time overhead introduced by BlueTracer. The results showed that the typical overhead per event is in the range of 40 - 220 microseconds per operation.
Then we firstly validated BlueTracer with Al-Khaser, an open-source application employed to assess how stealthy an analysis system is with respect to many evasion techniques used in the wild. BlueTracer was able to remain hidden to all the checks performed by Al-Khaser, thus proving to be effective at staying undetected. Furthermore, BlueTracer's tracking abilities also turned out to be efficacious, since evidence of the vast majority of Al-Khaser's techniques was present in BlueTracer's log. Lastly, we assessed BlueTracer's capabilities by conducting tests on real highly-evasive malware samples collected by Joe Security. The obtained results showed how BlueTracer was able to track the malicious activities performed by the malware instances, without them noticing it.

>>>>

In order to assess BlueTracer's effectiveness with respect to real malware instances, we tested it with five highly evasive samples collected by Joe Security \cite{JoeBox} having unusual evasion tricks that their specialized sandboxing products can take care of. For each sample, the evaluation of BlueTracer's tracing and stealth abilities was conducted by collecting BlueTracer's logs and checking if they were congruous with the sample's actions uncovered by Joe Sandbox, a commercial state-of-the-art stealthy analysis system \cite{JoeSandbox}. Moreover, we used Process Monitor, adequately cloaked by BluePill, as ground truth for what was carried out in the system.
We ran each sample for one hour using BlueTracer's \textbf{Main Image} mode.

@@

From the obtained results, we could deduce that BlueTracer was successful at tracing the samples' actions while remaining hidden, hence allowing them to perform their original malicious operations. An overview of the evasive actions executed from the samples' main executable which were tracked by BlueTracer and disabled by BluePill can be observed from \textit{Table 4.4} and \textit{Table 4.5}. In particular, the eight categories in which the actions are grouped correspond the ones adopted by BluePill. Specifically, the evasive patterns are classified as follows:

@@

In addition to counting the samples' events from their logs, we also recorded how many events were traced for each of eight different categories, namely, file activities, registry activities, process activities, thread activities, time activities, memory activities, system information-retrieval activities and user interface activities (\textit{Table 4.7}). Such groups are the same as the ones in Joe Sandbox's reports, which were, in fact, employed to determine the appropriate category for each event. From the data in \textit{Table 4.7} we can see that, for \textbf{1} and \textbf{3}, the most common operations are file activities (7.66\% and 59.68\%), coherently with the fact that one of their main actions is to drop files on the infected system. Regarding \textbf{2}, the largest portion of events involved thread activities (10.94\%) and memory activities (9.28\%), as we are dealing with a persistent dropper. Olympic, being a data-wiping malware, also mainly performs memory activities (9.44\%). Lastly, even for CCleaner, the majority of events are memory-related (19.84\%).

\section{Conclusions}
In this chapter we have outlined the results of BlueTracer's experimental evaluation. In order to assess BlueTracer's run-time overhead, we have tested it on a number of applications performing various tasks, employing a four different modes of operation. We have also introduced the overhead per event metric as a way of quantifying the run-time overhead introduced by BlueTracer. The results showed that the typical overhead per event is in the range of 40 - 220 microseconds per operation.
Then we firstly validated BlueTracer with Al-Khaser, an open-source application employed to assess how stealthy an analysis system is with respect to many evasion techniques used in the wild. BlueTracer was able to remain hidden to all the checks performed by Al-Khaser, thus proving to be effective at staying undetected. Furthermore, BlueTracer's tracking abilities also turned out to be efficacious, since evidence of the vast majority of Al-Khaser's techniques was present in BlueTracer's log. Lastly, we assessed BlueTracer's capabilities by conducting tests on real highly evasive malware samples collected by Joe Security. The obtained results showed how BlueTracer was able to track the malicious activities performed by the malware instances, without them noticing it.
